Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-05-31 VegOzone.tex
Refine and comments by Lisa

We determined the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees). Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations. The localization of parameters for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe. For a comprehensive list of MM parameters, consult Supplement~Table~S1. In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone-induced reduction of biomass when using MM Boreal parameterizations in ozone risk assessments.

@@

Initial simulations with the MM parameterizations for perennial grassland revealed an unexpected low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}). Substantial stomatal conductance occurred only during an extended warm period in late July. In ecological terms, this means that there would have been almost no growth of grass in the summer of 2019 -- a model prediction that is falsified by reality. We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland. This suggested that a localization of the parameterization to a subarctic climate was necessary.

To localize the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present). From these two subsets, we selected only dates within the GS (May--September) and plotted the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a). A comparison between the two PDFs indicates that the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$. It is unclear how fast the species composition of perennial grassland responds to these changes. Thus, the predominant acclimation at present is not known. Further, we overlaid these histograms with the MM $f_\mathrm{T}$. A low overlapping area suggests a low agreement with local climate conditions.

Hence, the idea was to define localized temperature response functions by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS. To this end, we defined two different hypothetical target species: subarctic and cold. We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures than MM, but slightly less efficient at warm temperatures than MM. This was accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ was kept at its MM value. In the same way, \emph{subarctic} was constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. The most extreme of the two localizations compared to MM is \emph{subarctic}.

@@

In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland. The MM parameterization is displayed as solid line.

@@

Similarly, we defined localized $f_\mathrm{T}$ and $f_\mathrm{light}$ for coniferous and deciduous trees. The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007}. Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}. 
To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}. From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$). For simplicity, the same interval was assumed for deciduous trees.

>>>>

We determined the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees). Basic parameters for these PFTs are derived from Boreal paramterizations \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations. The subarctic local climate derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe. For a comprehensive list of MM parameters, consult Supplement~Table~S1.

@@

Initial simulations with the Central European (Boreal is not available) MM parameterizations for perennial grassland revealed an unexpectedly low leaf level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}). Substantial stomatal conductance occurred only during an extended warm period in late July. In ecological terms, this means that there would have been almost no growth of grass in the summer of 2019 -- a model prediction that is falsified by reality. We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland. This suggested that deriving a local climate parameterization is necessary for appropriate modeling.

To derive a subarctic parameterization for the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present). From these two subsets, we selected only dates within the GS (May--September) and plotted the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a). A comparison between the two PDFs indicates that during the past decade the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$. It is unclear how fast the species composition or physiology via acclimation of perennial grassland responds to these changes. Thus, the predominant acclimation at present is not known. Further, we overlaid these histograms with the MM $f_\mathrm{T}$. A low overlapping area suggests a low agreement with local climate conditions.

We defined local climate derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS. We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species. This was accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ was kept at its MM value. In the same way, \emph{subarctic} was constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures. Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000. The most extreme of the two localizations compared to the MM is the \emph{subarctic}.

@@

In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland. As no Boreal parameterization is available for perennial grassland, the Central European MM parameterization is displayed as solid line.

@@

Similarly, we defined localized $f_\mathrm{T}$ and $f_\mathrm{light}$ for coniferous and deciduous trees. The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007}. Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year (\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}. 
To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}. From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$). For simplicity, the same interval was assumed for deciduous trees.
